--- 
title: "A Reproducible Research Compendium"
author: "cf. contributors"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This compendium is inteded is intended to be a growing knowledge base on reproducible research and by writing it collaboratively acts as learning-by-doing example on modern collaborative coding workflows and version control."
---

# Prerequisites

This is a _sample_ book written in **Markdown**. 
You can use anything that Pandoc's Markdown supports, e.g., 
a math equation $a^2 + b^2 = c^2$.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, 
and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install 
TinyTeX (which includes XeLaTeX): <https://yihui.name/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
